### 2023-01-22 manuscript.Rmd
update

---
title: "Towards Transparency and Open Science"
subtitle: "A Principled Perspective on Computational Reproducibility and Preregistration"
output:
  bookdown::pdf_document2:
    latex_engine: "xelatex"
    toc: true
    number_sections: false
    keep_tex:  true
header-includes: |
  \usepackage{libertine}
  %\usepackage{ebgaramond}
  %\usepackage{garamondlibre}
  %\usepackage{CormorantGaramond}
  %\usepackage{MinionPro}
  \usepackage{pdfpages}

  % below works well with hyperref, but spacing is standard (to large)
  %\usepackage{sectsty}
  %\sectionfont{\Large\normalfont\scshape\raggedright}
  %\subsectionfont{\large\normalfont\scshape\raggedright}

    % titlesec does not play nice with RMarkdown, so hack:
  \let\paragraph\oldparagraph
  \let\subparagraph\oldsubparagraph
  \usepackage[sc, compact, raggedright]{titlesec}
  % titlesec does not play nice with hyperref, so hack:
  \newcommand{\sectionbreak}{\phantomsection}
  \newcommand{\subsectionbreak}{\phantomsection}

  \usepackage{titling}
  \pretitle{
     \begin{minipage}{\textwidth}
     \includepdf[offset=5mm 0mm]{cover.pdf}
     \end{minipage}
     \newpage
     \clearpage%
     \thispagestyle{empty}%
     \addtocounter{page}{-2}%
     \null%
     \clearpage
     \begin{center}\LARGE}
  \posttitle{\end{center}}

  \usepackage{polyglossia}
  \setmainlanguage[]{english}
  \setotherlanguage[]{german}
fontsize: 11.5pt
linestretch: 1.2
# https://practicaltypography.com/page-margins.html
# 3.81 to 5.08cm or 45-90 characters or 2-3 alphabets
geometry: "left=4cm, right=4cm, top=3.5cm, bottom=4.5cm, bindingoffset=5mm"
classoption: twoside
papersize: a4
bibliography: references.bib
abstract: "`r paste0(c('\\\\setstretch{1.0}English:', readLines('abstract.md'), '\\\\par Deutsch:\\\\textgerman{', readLines('zusammenfassung.md'), '}\\\\newpage'), collapse = ' ')`"
csl: apa.csl
toc: false # it is manually placed
repro:
  files:
    - abstract.md
    - apa.csl
    - papers/
  tlmgr:
    - pgf
    - preview
    - libertine
    - pdfpages
  apt:
    - git
    - rsync
---

\newpage
\clearpage
\thispagestyle{empty}
\null
\clearpage

\begin{spacing}{1}
\tableofcontents
\end{spacing}

\newpage
\clearpage
\thispagestyle{empty}
\null
\clearpage

# Introduction

<!--psychology has unique difficulties-->
Psychology is a difficult science [@meehlTheoreticalRisksTabular1978].
Although there is some disagreement on why exactly this is the case, I doubt there is disagreement about the fact itself.
Laying open the difficulties and attempting to overcome them are not recent trends, though they have been invigorated by the so-called replication crisis in psychology [@ioannidisWhyMostPublished2005; @opensciencecollaborationEstimatingReproducibilityPsychological2015] that also begins to ripple through other empirical sciences.
As psychology grapples with this crisis of confidence in its empirical results, several causes and remedies have been proposed.
We can roughly divide the proposed countermeasures into two categories, those that aim to increase the correct use of statistical methods [e.g., @bakanTestSignificancePsychological20060327; @benjaminRedefineStatisticalSignificance2018; @cohenEarth051994;  @gigerenzerMindlessStatistics2004; @wagenmakersWhyPsychologistsMust2011] and those that aim to counteract sociological and psychological biases [e.g., @simmonsFalsepositivePsychologyUndisclosed2011; @rosenthalFileDrawerProblem1979; @bakkerRulesGameCalled2012; @johnMeasuringPrevalenceQuestionable2012].

<!--bla-transparency-bla-->
In my view, both categories try to address a lack of transparency about the inductive process in the empirical test of a theory.
Testing a theory empirically is often viewed as deductive, since the theory is making statements about the data.
Empirical scientists, however, often simultaneously engage in induction by deriving general statements from data.
I will argue that the inductive element in the empirical test leads to overconfidence in the empirical results, if unaccounted.
The extent of this bias depends on the inductive process.
The inductive process therefore has to be transparent to other researcher because they are otherwise unable to judge the empirical support of a theory.

The above distinction, between statistical and sociological countermeasures, arises from two different kinds of induction.
On the one hand, we have inductive processes that are well defined in the form of statistical methods, on the other hand researcher also engage in more informal inductive behavior outside of well defined models.
These different kinds of induction require different countermeasures.
Statistical methods make the inductive bias quantifiable; while open science measures reduce some uncertainty about the remaining informal sources of inductive bias.
To understand why transparency is crucial, we must understand that formal and informal induction are integral to empirical sciences.
Any science must be able to communicate how it generates its knowledge.
However, transparency has an outstanding role in psychology and other empirical sciences, because empirical statements lose their value without transparency about the inductive process that contributed to them.
Therefore, transparency is more than a virtue that may improve knowledge gains in empirical sciences but is an indispensable property.

<!--induction is necessary for psychological science-->
To function as an empirical science, psychology must be able to make statements about the world that can be compared to the actual conditions of the world.
In psychology, this is not a purely deductive endeavor.
Very few psychological theories are precise enough to derive testable statements.
While it is tempting to claim that a theory makes deductively testable claims by, e.g., implying a mean difference between two groups [@leePopperFalsificationCorroboration2021], such a statement is not testable on its own.
Consider a placebo and an experimental group, where we hypothesize and test a mean difference using a simple t-test.
The decision to reject the null hypothesis (groups have equal means) still depends, besides true mean difference and sample size, on the observed variance.
However, the variance needs to be estimated from the data.
So the threshold of the deductive decision depends on a quantity that must be induced.

<!--sometimes the use of induction is well justified-->
Induction is necessary and, in the present case, innocuous for psychology as a science.
Here, it is innocuous because the introduced bias can be accounted for and vanishes with increasing sample size, but we will see later that this is not always the case.
A t-test accounts for the estimation of variance by having somewhat wider tails than a z-test which assumes that the variance is known, i.e., the hurdle to reach significance is higher for a t-test (the critical value for a one sided test with α-level set to 1% and degrees of freedom set to 10, is t≈`r round(qt(.99, 10), 2)` vs. z≈`r round(qnorm(.99), 2)`).
It is widely known that even without this correction, the z-test is a good approximation when sample sizes are large [@studentProbableErrorMean1908].
Induction is also necessary because it is virtually impossible to ask psychological researchers to specify every detail, such as the variance, a priori from their theory.
If they had to, there would probably be no psychological theory that could be subjected to an empirical test.

<!--necessity of induction-->
In other words, induction gives our theories some slack to be imprecise and contain "blank" spaces, later to be filled based on observation.
It allows researchers to concentrate on the essential statements of their theories and choose some of the assumptions to fit the data well.
In some ways, it is the empirical researchers' answer to the Duhem–Quine problem [@duhemPhysicalTheoryExperiment1976; @vanormanquineTwoDogmasEmpiricism1976], which states that any empirical test of a theory is testing the conjunction of theory, auxiliary assumptions, and conjectures [@meehlTheoreticalRisksTabular1978; @meehlAppraisingAmendingTheories1990].
By inducing some quantities, psychological researchers can remove them from the conjunction.
If researchers use induction for some necessary but under the theory arbitrary assumptions, their theory will not be refuted because of these assumptions.
Since empirical researchers often cannot derive every assumption from their theory, avoiding refutation because of those assumptions is a desirable property.

<!--statistical induction can be dangerous-->
By the same token, whole theories may escape refutation by replacing every ill-fitting statement deduced from theory with statements induced from data.
If applied to auxiliary assumptions, such strategy of post hoc changing a theory in light of facts has been called "Lakatosian Defense" [@meehlAppraisingAmendingTheories1990].
If pushed to the limit, we arrive at a "theory" governed by the data.
Such a theory, full of empirically induced statements, is almost empty of statements that have been empirically verified.
The data used for induction can not refute these statements, so they they never have been subject of an empierical test.

<!--the appraisal of "induced theories" should be low-->
So what to think of such, yet untested, theory?
Researchers and philosophers of science differ considerably in their opinion about how to appraise theories, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p. 14] of a hypothesis.
Whatever measure they subscribe to, they would agree on a low appraisal of an untested theory.

<!--transparency about induction is required-->
So empirical researchers find themselves in a pinch.
On the one hand, they need induction to test their imprecise theories.
On the other hand, induction may render any test of a hypothesis ineffective.
The problem, I argue, therefore, is not induction but making transparent where and to what extend induction is used in the inferential process.
The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims as reported in the literature actually are.
Therefore, the question of this dissertation is *what* must be made transparent, and *how* to best make it transparent?

<!--forshadow the main part-->
The first question (the *what*) is theoretical in nature.
It is addressed in the synopses, which supplies the theoretical framework of the articles written as part of the dissertation.
Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge (sociological factors).

<!--how it relates to the articles-->
Based on this distinction, the articles answer the second question (the *how*).
I argue that transparency about statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
The conceptualization of computational reproducibility and preregistration as means for transparency is supplemented by practical guidance on how researchers may implement these approaches

# What necessitates transparency

The need for transparency is closely tied to the use of induction in the empirical test of a theory.
There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002, Ch.\ 5 Experience as a Method].
I do not attempt to rehash the debate about what constitutes an empirical test but aim to lay open the role of transparency in two frameworks that lend themselves to investigate induction.
The first framework motivates transparency when the aim of an empirical test is to evaluate the verisimilitude ("closeness to truth") of a theory.
The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.

Both frameworks show how unaccounted induction leads to overconfidence in empirical results and give us some theoretical tools to asses and control this overconfidence.
Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
These sections provide the basis to understand how computational reproducibility and preregistration enable a proper assessment of an empirical test on a conceptual level.

## An information-theoretic perspective

Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory [@niiniluotoVerisimilitudeThirdPeriod1998].
We can formalize the distance to the truth in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood to observe the state of the world $x$ where $f$ represents the ground truth.
We are now interested in how much information is lost if we use $g(x)$, our theory, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
<!--x is a single state, X is the random variable, and \mathcal{X} is all possible outcomes of X-->
\begin{align} 
\mathcal{L}_{KL}(f, g) &=\int_{\mathcal{X}}f(x)\log\left(f(x)\right)\mathrm{d}x - \int_{\mathcal{X}}f(x)\log\left(g(x)\right)\mathrm{d}x\\
&= \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]\label{eq:klexp}
\end{align}

Most readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch. 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
What is of interest here is not the derivation but how this conceptualization can help us to understand what happens when data is simultaneously used to induce quantities of a theory and test the theory.

This setup is, of course, highly theoretical.
$\mathcal{L}_{KL}(f, g)$ is unknowable, since the truth is unobserved.
However, this fact does not impede us from getting closer to the truth because we can still compare two theories relative to each other.
Because the expectation for $f$ remains constant (left-hand expectation in Eq.\ \@ref(eq:klexp)), we only need to estimate the relative expected loss of information (right-hand expectation in Eq.\ \@ref(eq:klexp)) to make a comparative judgment.
To make a relative judgment about several competing theories it suffices to estimate for any theory $g(x)$:
\begin{equation}
\mathbb{E}_{X \sim f}\left[\log(g(x))\right]
\end{equation}

To allow for quantities to be induced, we must assume that our theory is parameterized, e.g., $g(x|\theta)$.
That means our theory implies a family of possible probability distributions that may describe reality.
This parameterization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
Of those arbitrary assumptions, we want to find those that fit the reality with the least amount of information lost.
The best parameterization is achieved by:
\begin{equation}
\theta_* = \operatorname*{arg\,min}_\theta \mathcal{L}_{KL}(f, g(\cdot|\theta))
\end{equation}

The inference goal is for comparing other theories to $g(x)$ is, therefore:
\begin{equation}
\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right]
\end{equation}
Of course, we usually do not know $\theta_*$.
That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y$ are $n$ independent samples from $X\sim f$.

The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
It follows, almost surely, that we loose information:
\begin{equation}
\mathcal{L}_{KL}(f, g(\cdot|\hat{\theta}(y)) > \mathcal{L}_{KL}(f, g(\cdot|\theta_*))
\end{equation}
Or ignoring the constant:
\begin{equation}
\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right] >\mathbb{E}_{X \sim f}[\log(g(x|\hat{\theta}(y)))] 
\end{equation}

That is to say, any induced estimate will be sub-optimal.
The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
The point is to make a statement about the theory, not to make a statement about the data in combination with the theory.
If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased in the sense that it converges towards $\theta_*$, we may form an expectation over the data $Y$:
\begin{equation}
\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))]\label{eq:dataexp}
\end{equation}

Forming this expectation over data is a crucial step; it requires us to think beyond the data we observed to all the data we could have observed.
There are two ways to get at this expectation.
One is the use of Taylor series expansion, which follows in this section, and another is the use of cross-validation discussed in the next section.

We usually favor procedures to induce $\hat \theta$ that promise unbiased estimates for the observed likelihood, given that their assumptions are met.
The observed likelihood is, therefore, often available, e.g. in maximum likelihood estimation.
With slight abuse of notation, $\log(g(y|\theta)) \equiv \sum_{i = 1}^n \log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}\log(g(y|\hat\theta(y)))$ refers to the observed likelihood.

The expectation over the data together with Taylor series expansion yields:
\begin{equation}
\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] - tr[J(\theta_*)I(\theta_*)^{-1}]
\end{equation}

Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$ respectively.
For more details about this derivation see @burnhamModelSelectionMultimodel2002, Ch. 7.2.

The observed likelihood $\log(g(y|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
We may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
This overconfidence is directly related to how much induction a model entails.
This bias is often called the complexity or capacity of a model, i.e., how much the data is influencing the results and, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
I therefore denote it as as $\mathcal{C}$.
\begin{equation}
\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] + \mathcal{C}
\end{equation}

If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for the induction.
Fortunately, it is possible to approximate the complexity of a model under some conditions.
One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known [@boisbunonAkaikeInformationCriterion2014].

## A future performance perspective

In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
Instead of verisimilitude, one might be concerned with future performance.
That is, how well does a theory do in predicting novel facts.
Please note that the information-theoretic setup above has made no appeal to the expected performance on unseen data.
Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
In the future performance setup, we do not appeal to ground truth (I drop $\sim f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($\mathcal{L}$ stands for an arbitrary loss function):
\begin{equation}
\mathcal{L}(x, g(x|\theta)) = \mathbb{E}_xL(x, g(x|\theta))
\end{equation}

Again we can define the loss we have observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020]:
\begin{equation}
\mathbb{E}_y L(y, g(y|\hat\theta(y)))
 = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\theta))
\end{equation}

However, what we are interested in is not how well the theory did on data that informed it but on future, yet unseen, data:
\begin{equation}
\mathbb{E}_x\mathbb{E}_yL(x, g(x|\hat\theta(y)))
\end{equation}
This expectation over, what is often called training and test data, is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004].
Note that both training ($y$) and test ($x$) data vary in this expectation.
Therefore, it is closely related expectation over data shown in Eq.\ \@ref(eq:dataexp) in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].

Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
That is, we make use of cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
\begin{equation}
\mathbb{E}_{x}\mathbb{E}_{y}  L(x, g(x|\hat\theta(y)))
 = \frac{1}{n}\sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))
\end{equation}

As stated earlier, using cross-validation it is possible to  estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
To make the link to the first approach even more clear, we may define complexity in these terms [@hauensteinComputingAICBlackbox2018]:
\begin{equation}
\mathcal{C} = \frac{1}{n} \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i}))) - L(y_i, g(y_i|\hat\theta))
\end{equation}

Instead of a formal analysis to derive $\mathcal{C}$, we can simply repeat the inductive process, i.e., using cross-validation.
That drastically expands the set of inductive processes for which we can estimate the inductive bias since it is considerably easier to repeat the process than to derive the complexity analytically.

## A conceptual perspective

Now that we have established the need for transparency about the inductive process, we can drop a few more technical details to clarify what we have to make transparent.
It bears repeating that simply laying open what has been done is not enough.
Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.
On a conceptual level, we want to compare the following:
\begin{equation}
\mathcal{L}(\text{Theory}, \text{Reality})
\end{equation}

Where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
To allow for induction, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.

\begin{equation}
\mathcal{L}(\text{Model(Reality)}, \text{Reality})
\end{equation}
The idea is that we choose the version of our theory that best fits reality.
However, we are forced to rely on a limited sample of reality.
We are misled because these two factors, induction and limited sample size, interact.
Choosing the best version of our theory based on a sample is almost surely suboptimal.
Therefore, the observed loss is an overconfident estimate of the loss in the future and closeness to the truth.
\begin{equation}
\mathcal{L}(\text{Model(Reality)}, \text{Reality}) > \mathcal{L}(\text{Model(Sample)}, \text{Sample})
\end{equation}
\begin{equation}
\mathcal{L}(\text{Model(Reality)}, \text{Reality}) \approx \mathcal{L}(\text{Model(Sample)}, \text{Sample}) + \mathcal{C}(\text{Model})
\end{equation}

Transparency is necessary because induction leaves researchers overly optimistic regarding their theories' fit to the data ($\mathcal{C}$).
The extent of this optimism depends on the inductive process, not merely its results.
Specifically, it depends on the complexity, i.e., the ability of the inductive process to adapt to data.
The ability to adept to data is independent from the specific data that was observed.
Complexity is a function of the model, i.e., $\mathcal{C}(\text{Model})$, not of the data.
Without knowing the inductive process, researchers can not judge the overconfidence, so the inductive process ought to be made transparent.

# How to establish transparency

The above sections aimed to motivate the observation that the apparent fit of a theory to data is often overly optimistic if it has inductive elements.
This observation is only useful if we know the extent of this optimism.
However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{Model})$) and give us two starting points for making this bias transparent.
The first requires a formal analysis of the inductive process to compute the complexity (e.g., using information criteria), and the second merely requires that the process is repeatable (e.g., using cross-validation).
Both approaches require researchers to make the inductive process transparent instead of merely publishing the results.

Even a casual consideration of the above formalization should strike anyone who has ever worked with empirical data as unrealistic.
No one can expect researchers to be inductive only in formally analyzable or strictly repeatable ways.
The point is to set the goalpost and have a yardstick to measure how well we can judge empierical support.
Without induction, there is no bias to correct for.
With only formal induction, the bias can be quantified.
The issue of how to deal with informal induction remains.

As a first step, we can note that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
Formal, in the strict sense, means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
If a researcher employs a linear model, the complexity is calculable.
Even if it were not, we could fit the linear model on a large set of other data to asses its inductive behavior.
However, suppose the researcher reconsiders the model based on the results and their internal model.
For example, they decided based on a worse than expected results and their intuition that they should add a predictor to the model and remove another.
In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
We can not ask the researcher to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
However, we know that the resulting linear model is more complex than a linear regression formally implies.

$$
\mathcal{C}(\text{Model}) = \mathcal{C}(\text{formal}) +\mathcal{C}(\text{informal})
$$

That is not to say that we have no basis for judging the informal induction.
We can find the inductive decision reasonable and judge it unlikely that just about any variable would be added if the data suggests it.
Or the opposite, we might find it not well justified on theoretical grounds and deem it a purely data-driven decision, which implies higher complexity.
What one can say, however, is that this judgment is debatable and, therefore, subjective.

It is without question that researchers sometimes engage in inductive behavior that is neither formally analyzable nor repeatable.
This fact implies that for these situations, the complexity and hence the optimism bias can not be fully quantified.
The second best thing to transparency about complexity is to communicate the uncertainty about the complexity clearly.
However, judging the informal complexity $\mathcal{C}(\text{informal})$ must remain a subjective exercise.

The imperative to enable proper judgment of theories is simple: induce only what is necessary and what you induce should, if at all possible, be done formally.
Otherwise, the supposedly objective test of the theory using hard data must be judged more subjectively than necessary.

To summarise, I separate complexity $\mathcal{C}$ according to the transparency that can theoretically be achieved.
Formal inductive processes allow full transparency in the sense that $\mathcal{C}$ can be objectively quantified.
Informal inductive processes allow only limited transparency and must be subjectively judged.

Therefore, we have a theoretical bound that limits transparency about the inductive process.
Some things can not be made transparent in principle because some induction happens informally, so we can not estimate the complexity with certainty.
However, there is another bound that restricts transparency further.
That is how well is the inductive process, be it formal or informal, communicated to the intended recipient.
To achieve transparency about the inductive process, and ultimately about the empirical support of a theory, we must increase both bounds.

In the following, I propose preregistration as means to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bond).
I argue that both practices should ultimately provide transparency to judge empirical support for theories, though they may have secondary benefits for science.

## Transparency about statistical models: Computational\ Reproducibilibiltiy

Assume that a researcher only engaged in formal induction while testing a theory empirically.
If other researchers want to judge this empirical test properly, the inductive process must be communicated to them.
In other words, having formal induction is not enough, the induction still must be made transparent in practice.
Such transparency can be achieved by computational reproducibility.
Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992].
However, through the lens of transparency about the inductive process, we may further refine our view on computational reproducibility.

Consider software that delivers the same results upon input of the same data, i.e., it is reproducible.
However, place two restrictions on the software.
First, while able to compute results based on data, the software can not be understood by human researcher, e.g., because it is a binary program.
Second, when the data set changes, the software does not work anymore.
The first condition rules out that the software can be analyzed formally.
The second condition prevents us from changing the data and observing the results.
Therefore, we have no means to assess the complexity of the inductive process and can not judge the empirical support.
It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., we have to rule out one of the above conditions.

Suppose we relax the second condition, then we have a "black box" that is repeatable on other data.
In that case, we still have transparency about the inductive process in the sense that we can estimate the complexity of the inductive process, i.e., using cross-validation.
Relaxing the first condition, intransparency for human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
Therefore, we must extend the traditional definition of computational reproducibility and at least also require that the process can be repeated on other similar data.

We can, therefore, conclude that to provide transparency, computational reproducibility must satisfy two requirements.
First, computational reproducibility must ensure that the same data lead to the same results.
Second, computational reproducibility must enable the inductive process to be repeatable on similar data.

These two requirements are not easy to meet in practice.
In @peikertReproducibleDataAnalysis2021, I proposed a workflow that unifies both requirements under the objective of automating the full process from data to manuscript.
Implementing computational reproducibility by automating the process from data to results, fulfills both conditions needed for transparency while reducing the effort that must be invested.
Verifying that the results are actually produced by the inductive process and data is then a task that may be fulfilled by a computer without human intervention.
Removing the human from the loop facilitates that the data can be easily substituted by similar data.
Given enough computing power, automatic reproducibility allows to scale the inductive process to many data sets and therefore enables the judgment of the complexity.

While this workflow provides transparency and facilitates that the recipients of the work can judge the inductive process easily, it asks a lot from the researchers who create the work.
To address this problem and simplify the workflow's application in practice, I developed the R package `repro` [@R-repro].
I further refined and simplified the workflow to make it more accessible in @peikertReproducibleResearchTutorial2021 and contributed to @vanlissaWORCSWorkflowOpen2021's manuscript and the there presented software `worcs`.

To address the problem of automating reproducibility, these works divide it into four sub problems.
First, it must be unambiguous what results are generated by which inductive process.
This can be ensured by employing dynamic document creation.
Second, the version of the software that is used to generate the results must be known.
This can be ensured by documenting the version using software management.
Third, the version of the author written code (and possibly text) must be tracked using version control systems.
Fourth, how the data relates to software and computing infrastructure must be managed by e.g., workflow automation.

## Transparency about human researchers: Preregistration

Computational reproducibility enables transparency about formal inductive processes.
However, informal inductive processes deprive us of full transparency.
It is therefore prudent to replace informal induction with formal induction as much as possible.
If a researcher makes every effort to only employ formal inductive reasoning in a reproducible manner, they still must persuade their readers that they have not engaged in informal induction.

The problem is that formal inductive reasoning can be part of an informal process.
A perfectly reproducible linear regression is from the outside indistinguishable from one that was cherry picked from hundreds of possible regressions.
However, our expectation regarding verisimilitude and future performance between the cherry picked and the simple regression should vary substantially.

So what would constitute a persuasive argument that the data has not influenced the results above and beyond the complexity of the formal inductive process?
Researchers could simply try to explain the inductive process after the fact.
This is not an easy task, because we need to know what they did, as well as what they did not do, e.g., that they did not cherry pick, and what they would have done had the data looked different.
This goes back to the idea that simply knowing the outcome of an inductive process is not enough to judge the complexity.
To judge the complexity, the process must be transparent to the researcher, i.e., how would the results change if the data look different.

Trusting a post hoc explanation of an inductive process is not entirely unreasonable.
However, post hoc, formal and informal induction are indistinguishable unless the account is comprehensive.
Again, not only comprehensive for the data that was observed but also for all possible data.
This task is very difficult to accomplish in practice.
After all researchers would have to lay open the process that lead to the induction, i.e., their thoughts and their full mental model.
Dissolving the distinction between formal and informal induction would mean to surrender the advantage a formal inductive process provides; instead of quantification of complexity, the complexity must be judged with a considerable amount of uncertainty.

A clear distinction between formal and informal can be provided by specifying the formal inductive process before any data is available.
Such practice is called registration [@riceCurtailingUsePreregistration2019] or preregistration [@nosekPreregistrationRevolution2018]. 
Registration is a strong argument that the data has not influenced the results more than the complexity of the formal inductive process implies.
If the data does not yet exist, it can not influence the results.
Note that registration is about the flow of information, i.e., the data used for induction, and not about the temporal order.
Information can simply not travel backward in time, which makes registration before data acquisition so appealing.

Having no access to the data when specifying the analysis forces the researcher to think about the process of induction.
They have to reason about all the possible data patterns they might encounter.
While this makes registration so difficult, it is exactly the information we need and from which we may derive the complexity of the inductive process.

Note that requiring induction to be provably formal (analytically analyzable or repeatable) does not imply that induction itself is limited in its extend.
In other words, researchers still might be very unsure how the data will look and therefore rely on induction.
However, the must be clear what they are going to induce.
A formal inductive process can be designed to accommodate a wide range of data patterns, e.g., for a novel research question or a vague theory.
There are several strategies that reflect such uncertainties.
For example, if it is unknown which variables from a broad set of variables are involved, a variable selection mechanism can be specified or if it is unclear what functional form to expect, there exist many statistical models that allow almost arbitrary functional forms or if the theory does not suffice to define what an outlier is, this decision can made data dependent.

The idea of registering the inductive process without knowledge of the data to demarque formal from informal induction suffers from two problems.
First, how can this be accomplished in a way that is practical and fits into the well established practices of the scientific community?
Second, how to account for informal induction, e.g., when, despite all considerations, the data does not behave as anticipated or the registered method is deemed unsuitable because of other reasons?

In @peikertReproducibleResearchTutorial2021 I address the first problem and propose a practical, yet rigorous form of registration, termed preregistration as code (PAC).
In a PAC, researchers write the intended analyses as computer code, initially based on simulated data.
They include this code in a reproducible dynamic document written in the style of a traditional academic manuscript.
This version, with "mock" results based on simulated data serves as the registration.
When the data is gathered the results are updated to reflect the actual data.

To answer the second question of how to account for informal induction, I investigate the objective of registration and how it may be separated from the objective of confirmatory science in XXX.
Particularly, I propose to formalize the objective of registration as reduction of uncertainty about theoretical risk.
Being able to account for uncertainty is indispensable for the question of how to deal with informal induction.
If induction can not be repeated at will nor can it be formally analyzed, we are uncertain about the complexity of the inductive process.
Theoretical risk is conceptually related to the here discussed concept of complexity but put into the context of Bayesian philosophy of science.
Specifically, instead of the arbitrary loss functions considered here, the loss functions in XXX are restricted to those that apply to binary evidence and that satisfy the statistical relevancy condition (the loss function rewards the observation of evidence in favor of the theory, if the evidence is more likely under the theory).
For this class of loss functions, I show how to account for uncertainty caused by informal induction and that a reduction in uncertainty due to preregistration is universally beneficial.

# Discussion

This dissertation proposes transparency about the inductive process as an indispensable property of empirical sciences.
First, I address the question of *what* has to be made transparent.
The need for transparency is directly related to the use of induction in empirical sciences.
Induction is often necessary to test vague theories empirically.
However, induction introduces a bias that leads to overconfidence.
The extent of this overconfidence is a function of the inductive process.
Specifically, I propose that transparency must concern the extend to which data may influence the results and term this complexity.
I show that the complexity of an inductive process can be quantified, if the process can be analyzed or repeated on similar data.
I call such an inductive process "formal".
However, researchers often engage in informal induction, where complexity must be judged subjectively and with some uncertainty.
Second, I explain *how* computational reproducibility and preregistration provide transparency about complexity.
Preregistration allows to distinguish formal from informal induction, while computational reproducibility communicates the process of formal induction making it transparent.

This dissertation advances the theory and application of registration and computational reproducibility.
Based on theoretical consideration, I refined the requirements of computational reproducibility, so that not only the same data must produce the same results but also that the inductive process may be repeated on other similar data.
I developed a workflow for computational reproducibility that satisfies these requirements and build tooling to apply it.
Based on the computational reproducible workflow, I proposed preregistration as code (PAC) as an advancement of the practice of registration and develop the reduction of uncertainty about the theoretical risk as a formal objective of registration.

## From theorie to model

## Comparing predictions to reality

Subjectivity

## The issue of trust

## Summary

# Articles

In the following, I reprint the articles published as part of the dissertation as they were made accessible by the journals, as allowed by the respective licences.
Note, that the following articles were written collaboratively with the indicated coauthors.
The annex to § 6, para. 2 of the Doctoral Degree Regulations of the Faculty of Life Sciences, amended on 05.03.2015, *University Gazette of Humboldt-Universität zu Berlin 12* will be submitted with initiation of the doctoral degree procedure.

\newpage

## A Reproducible Data Analysis Workflow With R Markdown, Git, Make, and Docker

\phantomsection

The following article is reprinted from the following source under the conditions of the [Creative Commons Attribution (CC BY) 4.0 International License](https://creativecommons.org/licenses/by/4.0).

Peikert, A., & Brandmaier, A. M. (2021). A Reproducible Data Analysis Workflow With
R Markdown, Git, Make, and Docker. *Quantitative and Computational Methods in Behavioral Sciences*, *1*, e3763. https://doi.org/10.5964/qcmb.3763

\includepdf[pages=-, offset=5mm 0mm,frame=false]{papers/workflow.pdf}

## Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once

\phantomsection

The following article is reprinted from the following source under the conditions of the [Creative Commons Attribution (CC BY) 4.0 International License](https://creativecommons.org/licenses/by/4.0).

Peikert, A., van Lissa, C. J., & Brandmaier, A. M. (2021). Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once. *Psych*, *3*(4), 836–867. https://doi.org/10.3390/psych3040053

\includepdf[pages=-, offset=5mm 0mm]{papers/tutorial.pdf}

## Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

\phantomsection

The following article is prepared to be submitted to *Philosophical Psychology*.
The following preprint is reprinted from the following source under the conditions of the [Creative Commons Public Domain Dedication (CC0 1.0)](https://creativecommons.org/publicdomain/zero/1.0/).

\includepdf[pages=-, offset=5mm 0mm]{papers/prereg.pdf}

# References