### 2023-04-04 presentation.Rmd
add first version of talk slides

---
title: "Transparency in Science"
subtitle: "Principled Approaches to Computational Reproducibility and Preregistration"
author: 
  - "Aaron Peikert"
date: 'Defense | 22.05.2023'
output:
  xaringan::moon_reader:
    css:
    - xaringan-themer.css
    - "https://cdnjs.cloudflare.com/ajax/libs/animate.css/4.1.1/animate.min.css"
    nature:
      slideNumberFormat: |
        <div class="progress-bar-container">
          <div class="progress-bar" style="width: calc(%current% / %total% * 100%);">
          </div>
        </div>
      highlightStyle: solarized-dark
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
    seal: true
repro:
  packages:
    - xaringanthemer
    - showtext
    - svglite
    - patchwork
    - here
    - qrcode
---
class: center, middle

<style>
.remark-slide-number {
  position: inherit;
}

.remark-slide-number .progress-bar-container {
  position: absolute;
  bottom: 0;
  height: 2px;
  display: block;
  left: 0;
  right: 0;
}

.remark-slide-number .progress-bar {
  height: 100%;
  background-color: #b58900;
}

@keyframes khighlight {
  from {color: inherit;}
  to {color: var(--red);}
}

@keyframes ksuppress {
  from {color: inherit;}
  to {color: var(--grey);}
}

@keyframes kdehighlight {
  from {color: var(--red);}
  to {color: var(--textgrey);;}
}

/* The element to apply the animation to */
.highlight {
  animation-name: khighlight;
  animation-duration: 1s;
  animation-fill-mode: forwards;
}

.suppress {
  animation-name: ksuppress;
  animation-duration: 1s;
  animation-fill-mode: forwards;
}

.big-borders {
  margin-left: 150px;
  margin-right: 150px;
}

ul {
  list-style: none;
}

li::before {
  content: "⚬ ";
  color: var(--red);
}

.transparent {
  opacity: 0
}

</style>

```{r setup, include=FALSE}
library(xaringanthemer)
library(tidyverse)
library(patchwork)
library(qrcode)
options(htmltools.dir.version = FALSE)
knitr::read_chunk("R/plots.R")
knitr::opts_chunk$set(
  fig.width=9, fig.height=4, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = FALSE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE,
  dev = "svglite"
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
colors = list(
  red = "#dc322f",
  white = "#FFFFFF",
  black = "#002b36",
  grey = "#93a1a1",
  textgrey = "#657b83",
  yellow = "#b58900"
)

style_solarized_light(
  text_font_family = "LibertinusSansRegular",
  text_font_url = "https://fontlibrary.org//face/libertinus-sans",
  header_font_family = "LibertinusSerifDisplayRegular",
  header_font_url = "https://fontlibrary.org//face/libertinus-serif",
  text_font_size = "1.2rem",
  text_color = "#657b83",
  text_bold_color = "#b58900",
  text_slide_number_color = "#657b83",
  colors = unlist(colors)
)

style_extra_css(css = list(".todo" = list(color = "#17C25B"),
                           ".optional" = list(color = "#05BCE6"),
                           ".large" = list(`font-size` = "130%"),
                           ".huge" = list(`font-size` = "170%"),
                           ".Huge" = list(`font-size` = "250%"),
                           ".HUGE" = list(`font-size` = "600%"),
                           ".small" = list(`font-size` = "80%"),
                           ".enormous" = list(`font-size` = "1000%"),
                           ".full" = list(padding = "0px",
                                          margin = "0px",
                                          border = "0px"),
                           ".vertical" = list(`writing-mode` = "vertical-rl",
                                              `text-orientation` = "mixed"),
                           ".blurry" = list(color = "transparent",
                                            `text-shadow` = "0 0 5px rgba(0,0,0,0.5)"),
                           ".blurry2" = list(filter = "blur(2px)",
                                             `-webkit-filter` = "blur(2px)"),
                           ".curly" = list("font-family" = "Parisienne")))
```

# `r rmarkdown::metadata$title`

## `r rmarkdown::metadata$subtitle`

### `r rmarkdown::metadata$author`

### `r rmarkdown::metadata$date`

---
class: inverse, center, middle

# Trust in Science

---
class: center, middle

.enormous[?]

---
class: middle

```{r, echo=FALSE}
crisis <- tribble(
  ~response, ~bin, ~prop,
  "Yes, a signicant crisis", "YES", .52,
  "Yes, a slight crisis", "YES", .38,
  "No, there is no crisis", "NO", .03,
  "Don’t know", "NA", 0.07
)

crisis_plot <- crisis %>%
  group_by(bin) %>%
  summarise(prop = sum(prop)) %>%
  ungroup() %>% 
  ggplot(aes(
    "response",
    prop,
    fill = bin,
    label = bin,
    size = prop
  )) +
  geom_col() +
  geom_text(position = position_stack(vjust = 0.5), color = "#fdf6e3") +
  coord_flip() +
  theme_void() +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA)
  ) +
  scale_y_continuous(expand=c(0,0)) +
  scale_x_discrete(expand=c(0,0), limits = NULL) +
  scale_fill_manual(values = c("YES" = colors$red, "NO" = colors$black, "NA" = colors$grey)) +
  scale_size_continuous(range = c(4, 20)) +
  NULL
```

## Is there a replication crisis?*

.center[

```{r, out.width = '70%'}
crisis_plot /
  (crisis_plot +
     aes(label = str_c(prop * 100, "%")) + 
     scale_size_continuous(range = c(3, 12))) + 
  plot_layout(heights = c(.7, .3)) + 
  plot_annotation(theme = theme_void() + theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background = element_rect(fill = "transparent", colour = NA)
))

```
]

.small.right.animate__animated.animate__fadeIn.animate__slower[Baker, M. 1,500 scientists lift the lid on reproducibility. *Nature* 533, 452–454 (2016). https://doi.org/10.1038/533452a]

.small.right.animate__animated.animate__fadeIn.animate__slower[***They call it "reproducibility" and fail to publish the raw data.**]

---
class: center, middle

#.textgrey[Psychology is a difficult science]

.small.right.animate__animated.animate__fadeIn.animate__slower.animate__delay-2s[Meehl, P. E. (1978). Theoretical risks and tabular asterisks: Sir Karl, Sir Ronald, and the slow progress of soft psychology.  
*Journal of Consulting and Clinical Psychology*, 46(4), 806–834. https://doi.org/10.1037/0022-006X.46.4.806]

---
class: center, middle

.huge[Meehl, P. E. (1978). Theoretical risks and tabular asterisks:  
Sir Karl, Sir Ronald, and the slow progress of soft psychology.  
*Journal of Consulting and Clinical Psychology*, 46(4), 806–834.]

---
class: center, middle

.huge.supress[Meehl, P. E. (1978). Theoretical risks and tabular asterisks:  
Sir Karl, Sir Ronald, and the .highlight[slow progress] of .highlight[soft psychology].  
*Journal of Consulting and Clinical Psychology*, 46(4), 806–834.]

---
class:  middle, inverse, center

# The Progress of Soft Theories

---
class: center, middle

# Theories

.huge.animate__animated.animate__fadeIn.animate__slower[= allow to deduce predictions
]

---
class: center, middle

#.textgrey[Hard Theories]

.huge[= make .highlight[precise] predictions for given circumstances
]

---
class: center, middle

#.textgrey[Soft Theories]

.huge[= make .highlight[many] predictions for given circumstances
]

---
class: middle

.huge[

We allow predictions to be:

- close to truth
- probabilistic
- parameterized

]

---
class: center, middle

#.textgrey[Verisimilitude]

(=closeness to truth)

.huge[

$\mathcal{L}(\text{Theory}, \text{Data})$

]

---
class: center, middle

#.textgrey[Randomness]

(=probabilistic)

.huge[

$\mathcal{L}(\text{Theory}, \text{Data})$ is just an estimate.

]

---
class: center, middle

#.textgrey[Induction]

(=parameterized)

.huge[

$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$

]

---
class: middle, center, inverse

# Induction

.transparent[(=parameterized)]

.huge.transparent[

$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$

]

---
class: center, inverse

.huge[

$\text{Model}(\text{Theory}, \text{Data})$

]

Find the version of a theory, that is best supported by data.

--

or

Of any prediction that a theory can make, choose the one that minimizes $\mathcal{L}(\cdot, \text{Data})$.

--

or

```julia
for prediction in theory
  current = loss(prediction, data)
  if current < best
    best = current
  end
  return best
end
```

---
class: center, middle

.big-borders[

##.left.red[Induction]

###.textgrey[enables]

##.right.red[soft theories]

]

---
class: center, middle

.Huge[One]   .red[little]   .Huge[problem.]

---
class: center, middle

###.textgrey[Verisimilitude × Randomness × Induction]

---
class: center, middle

# Bias

---
class: center, middle

Psychology is difficult .transparent[→ soft theories → require induction → biased expectation → replication crisis]

---
class: center, middle

Psychology is difficult .animate__animated.animate__fadeIn.red[→ soft theories] .transparent[→ require induction → biased expectation → replication crisis]

---
class: center, middle

Psychology is difficult .dehighlight[→ soft theories] .animate__animated.animate__fadeIn.red[→ require induction] .transparent[→ biased expectation → replication crisis]

---
class: center, middle

Psychology is difficult → soft theories .dehighlight[→ require induction] .animate__animated.animate__fadeIn.red[→ biased expectation] .transparent[→ replication crisis]

---
class: center, middle

Psychology is difficult → soft theories → require induction .dehighlight[→ biased expectation] .animate__animated.animate__fadeIn.red[→ replication crisis]

---
class: center, middle

#.textgrey[

Why is .highlight[bias] a problem?

]

---
class: center, middle

#.blurry[Uncertainty]

---
class: center, middle

If the inductive bias is uncertain evaluating,

.huge[

$\mathcal{L}(\text{Theory}, \text{Data})$

]

becomes uncertain and unconvincing.

---
class: center, middle

$$\mathcal{L}(\text{Theory}, \text{Data})$$

$$=$$

$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$

$+$

.highlight[

$\mathcal{C}(\text{Model})$

]

.small.right.transparent[

This is similar to adjusted $R^2$, $C_p$, $AIC$, $BIC$, etc.

]

---
class: center, middle

$$\mathcal{L}(\text{Theory}, \text{Data})$$

$$=$$

$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$

$+$

.red[

$\mathcal{C}(\text{Model})$

]

.small.right[

This is similar to adjusted $R^2$, $C_p$, $AIC$, $BIC$, etc.

]

---
class: center, middle

.large.transparent[Under some circumstances,]

.huge.red[

$\mathcal{C}(\text{Model})$

]

.large[is the ability of a model to accommodate random data.]

---
class: center, middle

.large[Under some circumstances,]

.huge.red[

$\mathcal{C}(\text{Model})$

]

.large[can be approximated.]

---
class: center, middle

.large[ .highlight[Reproducibility] is a necessary* condition to estimate:]

.huge[
$\mathcal{C}(\text{Model})$]

.transparent.small.right.grey[*In an ideal world.]

---
class: center, middle

.large[

Some things are

.huge[.highlight[non]-reproducible]

on a fundamental level.

]

---
class: center, middle

.huge.red[

$\mathcal{C}(\text{Human})$

]

---
class: middle, center

.pull-left[

###.center[Reproducibility]

Allows estimating:

.large[

$$\mathcal{C}(\text{Statistical Model})$$

]

]

.pull-right[

###.center[Preregistration]

Does reduce:

.large[

$$\mathcal{C}(\text{Human})$$

]

]
---
class: center, middle

.large[.highlight[Reproducibility] is necessary to estimate:]

.large[

$\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model})$ .transparent[ $\require{cancel} + \cancel{\mathcal{C}(\text{Human})}$
]

]

.small.right.transparent[*In an ideal world.]

---
class: center, middle

.large[Reproducibility + .highlight[Preregistration] are sufficient* to estimate:]

.large[

$\require{cancel}\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model}) + \mathcal{C}(\text{Human})$

]

.small.right.grey[*In an ideal world.]

---
class: center, middle

.large[Reproducibility + .highlight[Preregistration] are sufficient* to estimate:]

.large[

$\require{cancel}\mathcal{L}(\text{Model}(\text{Theory}, \text{Data}), \text{Data})$ + $\mathcal{C}(\text{Model}) + \cancel{\mathcal{C}(\text{Human})}$

]

.small.right.grey[*In an ideal world.]
---
background-image: url("images/contribution1.svg")
background-position: center
background-size: contain
---
background-image: url("images/contribution2.svg")
background-position: center
background-size: contain
---
background-image: url("images/contribution3.svg")
background-position: center
background-size: contain
---
background-image: url("images/contribution4.svg")
background-position: center
background-size: contain
---
class: middle

**.red[1\. Paper:]**  
Peikert, A., & Brandmaier, A. M. (2021). A Reproducible Data Analysis Workflow. *Quantitative and Computational Methods in Behavioral Sciences*, *1*. https://doi.org/10.5964/qcmb.3763

**2\. Paper:**  
Peikert, A., van Lissa, C. J., & Brandmaier, A. M. (2021). Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once. *Psych*, *3*(4), 836–867. https://doi.org/10.3390/psych3040053

**3\. Paper:**  
Peikert, A., & Brandmaier, A. M. (2023). *Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.* PsyArXiv. .textgrey[https://doi.org/10.31234/osf.io/cs8wb]

**Synopsis:**  
Peikert, A. (2023). *Towards Transparency and Open Science: A Principled Perspective on Computational Reproducibility and Preregistration* [Doctoral dissertation, Humboldt-Universität zu Berlin]. Zenodo. https://www.doi.org/10.5281/zenodo.7654989

---
class: inverse, middle, center

# Automate Reproducibility

---
class:  middle, center

# Four Problems

---
class:  middle, center

# Four Concepts

---
class:  middle, center

# Four Tools

---
class: middle

.pull-left.large[
.textgrey[1\. Copy&Paste Mistakes]  
.transparent[2\. Conflicting Versions]  
.transparent[3\. Missing Software]  
.transparent[4\. Unclear Procedure]  
]

.pull-right.large[
.transparent[Dynamic Documents]  
.transparent[Version Control]  
.transparent[Software Management]  
.transparent[Workflow Orchestration]  
]
---
class: middle

.pull-left.large[
.textgrey[1\. Copy&Paste Mistakes]  
.transparent[2\. Conflicting Versions]  
.transparent[3\. Missing Software]  
.transparent[4\. Unclear Procedure]  
]

.pull-right.large[
.highlight[Dynamic Documents]  
.transparent[Version Control]  
.transparent[Software Management]  
.transparent[Workflow Orchestration]  
]
---
class: middle

.pull-left.large[
.suppress[1\. Copy&Paste Mistakes]  
.textgrey[2\. Conflicting Versions]  
.transparent[3\. Missing Software]  
.transparent[4\. Unclear Procedure]  
]

.pull-right.large[
.suppress[Dynamic Documents]  
.transparent[Version Control]  
.transparent[Software Management]  
.transparent[Workflow Orchestration]  
]
---
class: middle

.pull-left.large[
.grey[1\. Copy&Paste Mistakes]  
.textgrey[2\. Conflicting Versions]  
.transparent[3\. Missing Software]  
.transparent[4\. Unclear Procedure]  
]

.pull-right.large[
.grey[Dynamic Documents]  
.highlight[Version Control]  
.transparent[Software Management]  
.transparent[Workflow Orchestration]  
]
---
class: middle

.pull-left.large[
.grey[1\. Copy&Paste Mistakes]  
.suppress[2\. Conflicting Versions]  
.textgrey[3\. Missing Software]  
.transparent[4\. Unclear Procedure]  
]

.pull-right.large[
.grey[Dynamic Documents]  
.suppress[Version Control]  
.transparent[Software Management]  
.transparent[Workflow Orchestration]  
]
---
class: middle

.pull-left.large[
.grey[1\. Copy&Paste Mistakes]  
.grey[2\. Conflicting Versions]  
.textgrey[3\. Missing Software]  
.transparent[4\. Unclear Procedure]  
]

.pull-right.large[
.grey[Dynamic Documents]  
.grey[Version Control]  
.highlight[Software Management]  
.transparent[Workflow Orchestration]  
]
---
class: middle

.pull-left.large[
.grey[1\. Copy&Paste Mistakes]  
.grey[2\. Conflicting Versions]  
.suppress[3\. Missing Software]  
.textgrey[4\. Unclear Procedure]  
]

.pull-right.large[
.grey[Dynamic Documents]  
.grey[Version Control]  
.suppress[Software Management]  
.transparent[Workflow Orchestration]  
]
---
class: middle

.pull-left.large[
.grey[1\. Copy&Paste Mistakes]  
.grey[2\. Conflicting Versions]  
.grey[3\. Missing Software]  
.textgrey[4\. Unclear Procedure]  
]

.pull-right.large[
.grey[Dynamic Documents]  
.grey[Version Control]  
.grey[Software Management]  
.highlight[Workflow Orchestration]  
]

---
class: middle

**1\. Paper:**  
Peikert, A., & Brandmaier, A. M. (2021). A Reproducible Data Analysis Workflow. *Quantitative and Computational Methods in Behavioral Sciences*, *1*. https://doi.org/10.5964/qcmb.3763

**.red[2\. Paper:]**  
Peikert, A., van Lissa, C. J., & Brandmaier, A. M. (2021). Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once. *Psych*, *3*(4), 836–867. https://doi.org/10.3390/psych3040053

**3\. Paper:**  
Peikert, A., & Brandmaier, A. M. (2023). *Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.* PsyArXiv. .textgrey[https://doi.org/10.31234/osf.io/cs8wb]

**Synopsis:**  
Peikert, A. (2023). *Towards Transparency and Open Science: A Principled Perspective on Computational Reproducibility and Preregistration* [Doctoral dissertation, Humboldt-Universität zu Berlin]. Zenodo. https://www.doi.org/10.5281/zenodo.7654989

---
class: inverse, middle, center

# Reproducibility × Preregistration

---
class: center, middle

.pull-left[

### Standard Preregistration

hunches

↓

preregistration

↓

data

↓

article draft

]

--

.pull-right[

### Preregistration as Code

simulated data

↓

article draft with mock results

↓

data

↓

article draft with real results

]
---
class: middle

**1\. Paper:**  
Peikert, A., & Brandmaier, A. M. (2021). A Reproducible Data Analysis Workflow. *Quantitative and Computational Methods in Behavioral Sciences*, *1*. https://doi.org/10.5964/qcmb.3763

**2\. Paper:**  
Peikert, A., van Lissa, C. J., & Brandmaier, A. M. (2021). Reproducible Research in R: A Tutorial on How to Do the Same Thing More Than Once. *Psych*, *3*(4), 836–867. https://doi.org/10.3390/psych3040053

**.red[3\. Paper:]**  
Peikert, A., & Brandmaier, A. M. (2023). *Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.* PsyArXiv. .textgrey[https://doi.org/10.31234/osf.io/cs8wb]

**Synopsis:**  
Peikert, A. (2023). *Towards Transparency and Open Science: A Principled Perspective on Computational Reproducibility and Preregistration* [Doctoral dissertation, Humboldt-Universität zu Berlin]. Zenodo. https://www.doi.org/10.5281/zenodo.7654989

---
class: inverse, middle, center

# The Goal of Preregistration

---
class: middle, center

#.blurry2[Uncertainty]

---
background-image: url("images/contribution.svg")
background-position: center
background-size: contain
---